Toadflax Analysis

### 2024-03-08 toadflax_report.qmd
3/8 first commit

```{r scores plot, echo = F}
ggplot() +
  geom_point(data = model.data.all,
             aes(PC1, PC2, color = type)) +
  # geom_text(aes(label = day))
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
  labs(x = "PC1", y = "PC2", 
       title = "Toadflax PCA Plot",
       color = element_blank()) +
  stat_ellipse(data = model.data.all,
               aes(PC1, PC2, color = type)) +
  labs(title = "Toadflax PCA Scores Plot") +
  theme_bw() +
  scale_color_manual(values = c("DT" = "#F8766D", 
                                "Synthetic" = "#7CAE00",
                                "YT" = "#00BFC4",
                                "Natural" = "#C77CFF"))

@@

```{r loadings plot, echo = F}
ggplot() +
  geom_segment(data = loadings.plot,
               aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.1, "inches")),
               alpha = 0.5,
               color = "red") +
  geom_text(data = loadings.plot,
            hjust = 1.1, vjust = -1.2,
            aes(PC1, PC2, label = compound),
            size = 4) +
  labs(x = "PC1", y = "PC2", title = "Toadflax PCA Loadings") +
  theme_bw() +
  scale_x_continuous(expand = c(0.1, 0.1)) +
  scale_y_continuous(expand = c(0.1, 0.1))
```

@@

```{r sc.lo.plot, echo = F}
ggbiplot(pca.results, obs.scale = 1, var.scale = 1, ellipse = FALSE,
         group = data.clr$type, circle = FALSE) +
  theme_bw() +
  labs(title = "Toadflax PCA with loadings and scores")
```

@@

```{r scores plot category POST natural removed, echo = F, fig.width = 12}
all.scores <-
  ggplot() +
    geom_point(data = model.data.all,
               aes(PC1, PC2, color = type)) +
    # geom_text(aes(label = day))
    geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
    geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
    labs(x = "PC1", y = "PC2", 
         title = "Toadflax PCA Plot",
         color = element_blank()) +
    stat_ellipse(data = model.data.all,
                 aes(PC1, PC2, color = type)) +
    labs(title = "Toadflax PCA - All Data") +
    theme_bw() +
    scale_color_manual(values = c("DT" = "#F8766D", 
                                  "Synthetic" = "#7CAE00",
                                  "YT" = "#00BFC4",
                                  "Natural" = "#C77CFF"))

scores.post.natrem <- 
  ggplot() +
  geom_point(data = model.data.all %>%
               filter(!type == "Natural"),
             aes(PC1, PC2, color = type)) +
  # geom_text(aes(label = day))
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
  labs(x = "PC1", y = "PC2", 
       title = "Toadflax PCA Plot",
       color = element_blank()) +
  stat_ellipse(data = model.data.all %>%
               filter(!type == "Natural"),
               aes(PC1, PC2, color = type)) +
  labs(title = "Toadflax PCA - DT, YT, Synthetic") +
  theme_bw() +
  # theme(legend.position = "none") +
  scale_color_manual(values = c("DT" = "#F8766D", 
                                "Synthetic" = "#7CAE00",
                                "YT" = "#00BFC4",
                                "Natural" = "#C77CFF"))

grid.arrange(scores.post.natrem, all.scores, ncol = 2)

@@

grid.arrange(scores.pre.natrm, all.scores, ncol = 2)
```

@@

grid.arrange(scores.post.hybrids, all.scores, ncol = 2)

@@

grid.arrange(scores.pre.hybrids, all.scores, ncol = 2)
```

@@

```{r PCA plant, echo = F}
pca.results <-
data.clr %>%
  dplyr::select(-c(plant_id, plant, type, day)) %>%
  prcomp(scale = T)

@@

```{r scores plot plant, echo = F}
ggplot() +
  geom_point(data = model.data,
             aes(PC1, PC2, color = plant)) +
  # geom_text(aes(label = day))
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
  labs(x = "PC1", y = "PC2", 
       title = "Toadflax PCA Plot",
       color = element_blank()) +
  stat_ellipse(data = model.data,
               aes(PC1, PC2, color = plant)) +
  labs(title = bquote("Toadflax PCA Scores Plot by "*bold("Plant"))) +
  theme_bw() 
```

>>>>

```{r scores plot, echo = F}
scores.all <-
ggplot() +
  geom_point(data = model.data.all,
             aes(PC1, PC2, color = type)) +
  # geom_text(aes(label = day))
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
  labs(x = "PC1", y = "PC2", 
       title = "Toadflax PCA Scores Plot") +
  stat_ellipse(data = model.data.all,
               aes(PC1, PC2, color = type)) +
  theme_bw() +
  scale_color_manual(values = c("DT" = "#F8766D", 
                                "Synthetic" = "#7CAE00",
                                "YT" = "#00BFC4",
                                "Natural" = "#C77CFF"))

ggplotly(scores.all,
         width = 700) %>%
  layout(legend = list(title = "")) %>%
  config(
    displayModeBar = TRUE,
    displaylogo = FALSE,
    modeBarButtonsToRemove = list('lasso2d', 
                                  'hoverClosestCartesian',
                                  'hoverCompareCartesian')
  )
```

@@

```{r loadings plot, echo = F}
loadings.ggplot <-
ggplot() +
  geom_segment(data = loadings.plot,
               aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.1, "inches")),
               alpha = 0.5,
               color = "red") +
  geom_text(data = loadings.plot,
            hjust = 1.1, vjust = -1.2,
            aes(PC1, PC2, label = compound),
            size = 4) +
  labs(x = "PC1", y = "PC2", title = "Toadflax PCA Loadings") +
  theme_bw() +
  scale_x_continuous(expand = c(0.1, 0.1)) +
  scale_y_continuous(expand = c(0.1, 0.1))

loadings.plotly <-
ggplotly(loadings.ggplot,
         width = 700) %>%
  layout(legend = list(title = "")) %>%
  config(
    displayModeBar = TRUE,
    displaylogo = FALSE,
    modeBarButtonsToRemove = list('lasso2d', 
                                  'hoverClosestCartesian',
                                  'hoverCompareCartesian')
  )

loadings.plotly
```

@@

```{r sc.lo.plot, echo = F}
ggplotly(
ggbiplot(pca.results, obs.scale = 1, var.scale = 1, ellipse = FALSE,
         group = data.clr$type, circle = FALSE) +
  theme_bw() +
  labs(title = "Toadflax PCA with loadings and scores"),
width = 700) %>%
  layout(legend = list(title = "")) %>%
  config(
    displayModeBar = TRUE,
    displaylogo = FALSE,
    modeBarButtonsToRemove = list('lasso2d', 
                                  'hoverClosestCartesian',
                                  'hoverCompareCartesian')
)
```

@@

```{r scores plot category POST natural removed, echo = F, fig.width = 12}
# scores.all.plotly <-
# ggplotly(
scores.all <-
  ggplot() +
    geom_point(data = model.data.all,
               aes(PC1, PC2, color = type)) +
    # geom_text(aes(label = day))
    geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
    geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
    labs(x = "PC1", y = "PC2", 
         title = "Toadflax PCA Plot",
         color = element_blank()) +
    stat_ellipse(data = model.data.all,
                 aes(PC1, PC2, color = type)) +
    labs(title = "Toadflax PCA - All Data") +
    theme_bw() +
    scale_color_manual(values = c("DT" = "#F8766D", 
                                  "Synthetic" = "#7CAE00",
                                  "YT" = "#00BFC4",
                                  "Natural" = "#C77CFF"))

scores.post.natrem <-
  ggplot() +
  geom_point(data = model.data.all %>%
               filter(!type == "Natural"),
             aes(PC1, PC2, color = type)) +
  # geom_text(aes(label = day))
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
  labs(x = "PC1", y = "PC2", 
       title = "Toadflax PCA Plot",
       color = element_blank(),
       color = element_blank()) +
  stat_ellipse(data = model.data.all %>%
               filter(!type == "Natural"),
               aes(PC1, PC2, color = type)) +
  labs(title = "Toadflax PCA - DT, YT, Synthetic") +
  theme_bw() +
  # theme(legend.position = "none") +
  scale_color_manual(values = c("DT" = "#F8766D", 
                                "Synthetic" = "#7CAE00",
                                "YT" = "#00BFC4",
                                "Natural" = "#C77CFF"))

# remove the legend from one of the plots
# for (i in 1:length(scores.all.plotly$x$data)) {
#   scores.all.plotly$x$data[[i]]$showlegend <- FALSE
# }
# use subplot to plot the two plots side by side
# subplot(scores.all.plotly, scores.post.natrem, 
#         nrows = 1, shareY = FALSE, shareX = FALSE, 
#         widths = c(0.5, 0.5)) %>%
#   layout(legend = list(title = ""))

grid.arrange(scores.post.natrem, scores.all, ncol = 2)

@@

grid.arrange(scores.pre.natrm, scores.all, ncol = 2)
```

@@

grid.arrange(scores.post.hybrids, scores.all, ncol = 2)

@@

grid.arrange(scores.pre.hybrids, scores.all, ncol = 2)
```

@@

### All Data
```{r PCA plant, echo = F}
pca.results <-
data.clr %>%
  dplyr::select(-c(plant_id, plant, type, day)) %>%
  prcomp(scale = T)

@@

```{r scores plot plant, echo = F}
scores.plant <-
ggplot() +
  geom_point(data = model.data,
             aes(PC1, PC2, color = plant),
             alpha = 0.8) +
  # geom_text(aes(label = day))
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
  labs(x = "PC1", y = "PC2", 
       title = "Toadflax PCA Plot",
       color = element_blank()) +
  stat_ellipse(data = model.data,
               aes(PC1, PC2, color = plant)) +
  labs(title = bquote("Toadflax PCA Scores Plot by "*bold("Plant"))) +
  scale_color_manual(values = c("HBR" = "#F8766D", 
                                "RAD" = "#7CAE00",
                                "D6" = "#00BA38",
                                "Y6" = "#00BFC4",
                                "Y6D6" = "#619CFF",
                                "D6Y6" = "#F564E3")) +
  theme_bw()

scores.plant
```

### Variable Plant Selection

#### Compare: HBR and RAD

##### Post-PCA Model Filter
```{r scores plot type Post HBR RAD, echo = F, fig.width = 12}
scores.post.hbrrad <-
ggplot() +
  geom_point(data = model.data.all %>%
               filter(plant %in% c("HBR", "RAD")),
             aes(PC1, PC2, color = plant),
             alpha = 0.8) +
  # geom_text(aes(label = day))
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
  labs(x = "PC1", y = "PC2", 
       title = "Toadflax PCA - HBR and RAD",
       color = element_blank()) +
  stat_ellipse(data = model.data.all %>%
               filter(plant %in% c("HBR", "RAD")),
               aes(PC1, PC2, color = plant)) +
  # geom_polygon(data = model.data.all %>%
  #                 filter(plant %in% c("HBR", "RAD")),
  #               aes(x = PC1, y = PC2, fill = plant),
  #               alpha = 0.1, color = NA,
  #               stat = "ellipse") +
  theme_bw() +
  scale_color_manual(values = c("HBR" = "#F8766D", 
                                "RAD" = "#7CAE00",
                                "D6" = "#00BA38",
                                "Y6" = "#00BFC4",
                                "Y6D6" = "#619CFF",
                                "D6Y6" = "#F564E3"))

grid.arrange(scores.post.hbrrad, scores.plant, ncol = 2)
```

##### Pre-PCA Model Filter
```{r PCA category PRE hbr rad, echo = F}
plants <- c("RAD", "HBR")

pca.results <-
data.clr %>%
  filter(plant %in% plants) %>%
  dplyr::select(-c(plant_id, plant, type, day)) %>%
  prcomp(scale = T)

data.plus.pca <-
  data.clr %>%
    filter(plant %in% plants) %>%
    cbind(scores)

# meta data
meta <- data.clr %>% 
  filter(plant %in% plants) %>%
  dplyr::select(plant, type, day)

model.data.wild <- data.frame(PC1, PC2, meta)
model.data$day <- factor(model.data$day, levels = c("1", "2", "3", "4",
                                                    "5", "6", "7", "8",
                                                    "9"))

```
```{r scores plot type Pre HBR RAD, echo = F, fig.width = 12}
scores.post.hbrrad <-
ggplot() +
  geom_point(data = model.data.wild,
             aes(PC1, PC2, color = plant),
             alpha = 0.8) +
  # geom_text(aes(label = day))
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
  labs(x = "PC1", y = "PC2", 
       title = "Toadflax PCA - HBR and RAD",
       color = element_blank()) +
  stat_ellipse(data = model.data.wild,
               aes(PC1, PC2, color = plant)) +
  # geom_polygon(data = model.data.all %>%
  #                 filter(plant %in% c("HBR", "RAD")),
  #               aes(x = PC1, y = PC2, fill = plant),
  #               alpha = 0.1, color = NA,
  #               stat = "ellipse") +
  theme_bw() +
  scale_color_manual(values = c("HBR" = "#F8766D", 
                                "RAD" = "#7CAE00",
                                "D6" = "#00BA38",
                                "Y6" = "#00BFC4",
                                "Y6D6" = "#619CFF",
                                "D6Y6" = "#F564E3"))

grid.arrange(scores.post.hbrrad, scores.plant, ncol = 2)
```
